---
title: "cellchat: skin versus RB neurons"
authors: "Nick Calistri (calistri@ohsu.edu) & Alex Nechiporuk" 
date: "2024-04-02"
output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: true
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up


## Load libraries
```{r}
library(tidyverse)
library(Seurat)
library(CellChat)
```

 
 
 

## Load data

```{r}
pLLg <- readRDS('30h_pioneers_followers.rds')

pLLP <- readRDS('pLLP.rds')

```



## Visualize Suerat object

```{r}
DimPlot(pLLg,
        label = TRUE)+
  coord_equal()
FeaturePlot(pLLg, features = c('hoxb5b', "ret"), order = T,
            min = 0.1,
            pt.size = 1)+ NoAxes()+coord_equal()


DimPlot(pLLP,
        label = TRUE)+
  coord_equal()

FeaturePlot(pLLP, features = c('hmx2'), order = T,
            min = 0.1,
            pt.size = 1)+ NoAxes()+coord_equal()


```


## Adjust metadata and convert to dietSeurat

```{r}
pLLg@meta.data$celltype <- paste0('pLLg_', Idents(pLLg))
pLLP@meta.data$celltype <- paste0('pLLP_', Idents(pLLP))

DefaultAssay(pLLg) <- 'RNA'

ds_pLLg <- DietSeurat(pLLg,
                      assays = 'RNA')
ds_pLLP <- DietSeurat(pLLP,
                      assays = 'RNA')
```


##Merge objects
```{r}
data.combined <- merge(x = ds_pLLg, y = ds_pLLP, add.cell.ids = c("pLLg", "pLLP"), project = "data.combined")

data.combined[['condition']] <- str_split(data.combined@meta.data$orig.ident, pattern = '_', simplify = TRUE)[,1]
head(colnames(data.combined))
table(data.combined$orig.ident)

data.combined.meta <- data.combined@meta.data

Idents(data.combined) <- 'celltype'
```


```{r}
data.combined[["percent.mt"]] <- PercentageFeatureSet(data.combined, pattern = "^mt-|^COX3$")
# Visualize QC metrics as a violin plot
VlnPlot(data.combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0)

```

## normalize data with SCTransform

```{r}
data.combined <- SCTransform(data.combined)

# Note: 08/19/24 NLC - Commenting out this section and switching to using SCTransform due to read-depth disparity between pLLg and pLLP data sets

# data.combined <- NormalizeData(data.combined)
# data.combined <- FindVariableFeatures(data.combined, selection.method = 'vst', nfeatures = 2000)
# top10_variablefeatures <- head(VariableFeatures(data.combined), 10)
# 
# var_feature_plot <- VariableFeaturePlot(data.combined)
# var_feature_plot <- LabelPoints(plot = var_feature_plot, points = top10_variablefeatures, repel = TRUE)
# 
# data.combined <- ScaleData(data.combined, features = rownames(data.combined))

```

```{r}
data.combined <- RunPCA(data.combined)
VizDimLoadings(data.combined, dims = 1:4, reduction = 'pca')
ElbowPlot(data.combined, ndims = 50)+
  geom_vline(xintercept = 10.5)
```

```{r}
data.combined <- FindNeighbors(data.combined, dims = 1:10)
data.combined <- FindClusters(data.combined, resolution = .1)
data.combined <- RunUMAP(data.combined, dims = 1:10)

pLLg_pLLP_plot <- DimPlot(data.combined, label = F, label.size = 5, group.by = "celltype")+
  coord_equal()
ggsave(pLLg_pLLP_plot,  filename =  "pLLg_pLLP_plot.jpeg", dpi = 600, scale = 0.25)


data.combined.meta <- data.combined@meta.data
#saveRDS(data.combined, "data.combined.rds")
```

##
```{r}

hmx2 <- FeaturePlot(data.combined, features = c("hmx2"), order = T, 
            min = 0.1,
            pt.size = 1)+ NoAxes()+coord_equal()
ggsave(hmx2,  filename =  "hmx2.jpeg", dpi = 600, scale = 0.25)

ret <- FeaturePlot(data.combined, features = c('ret'), order = T,
            min = 0.1,
            pt.size = 1)+ NoAxes()+coord_equal()
ggsave(ret,  filename =  "ret.jpeg", dpi = 600, scale = 0.25)

nr2f2 <- FeaturePlot(data.combined, features = c('nr2f2'), order = T,
            min = 0.1,
            pt.size = 1)+ NoAxes()+coord_equal()
ggsave(nr2f2,  filename =  "nr2f2.jpeg", dpi = 600, scale = 0.25)


```


## Convert to cellchat

```{r}
# Extract relevant info from seurat object
data.input <- data.combined[["SCT"]]@data # normalized data matrix
# For Seurat version >= “5.0.0”, get the normalized data via `seurat_object[["RNA"]]$data`
meta <- data.frame(group = data.combined@meta.data$celltype,
                   row.names = row.names(data.combined@meta.data$celltype))

# Create cellchat object
cellchat <- createCellChat(object = data.input,
                           meta = meta,
                           group.by = 'group')


# Add meta data
cellchat <- addMeta(cellchat, meta = meta, meta.name = 'celltype')
setIdent(cellchat, ident.use = 'celltype')
levels(cellchat@idents)

```

## 24/12/1: Update cellchat ligand/receptor dictionary with Teneurin-latrophilin interactions


```{r}
CellChatDB <- CellChatDB.zebrafish

showDatabaseCategory(CellChatDB)

cellchat@DB <- CellChatDB
```

```{r}
# Read ligand-receptor data for teneurin-latrophilin interactions
ten_lat_interaction <- read_csv('teneurin_latrophilin_interaction.csv') %>%
  as.data.frame()

row.names(ten_lat_interaction) <- ten_lat_interaction$interaction_name

ten_lat_geneInfo <- read_csv('teneurin_latrophilin_geneInfo.csv') %>%
  as.data.frame()

row.names(ten_lat_geneInfo) <- ten_lat_geneInfo$Symbol

# Copy the cellchatDB
CellChatDB2 <- CellChatDB.zebrafish

# Combine with regular zebrafish ligand-receptor interaction dictionary
CellChatDB2$interaction <- rbind(CellChatDB2$interaction,
                                ten_lat_interaction)

CellChatDB2$geneInfo <- rbind(CellChatDB2$geneInfo,
                                ten_lat_geneInfo)

cellchat@DB <- CellChatDB2
```


## Preprocess the expression data (does not project via ppi due to missing zebrafish ppi)

```{r}
cellchat <- subsetData(cellchat)

future::plan('multisession', workers = 6)

```

## Compute communication probability

```{r}
if(file.exists('241201_cellchat_pLLg_pLLP.rds')){
  print('Loading saved cellchat object')
  cellchat <- readRDS('241201_cellchat_pLLg_pLLP.rds')
}else{
  print('Computing cellchat interactions, ETA: ~15min')
  # Identify overexpressed genes and interactions
  cellchat <- identifyOverExpressedGenes(cellchat)
  cellchat <- identifyOverExpressedInteractions(cellchat)
  
  # Compute communication probability
  cellchat <- computeCommunProb(cellchat, raw.use = TRUE)
  cellchat <- filterCommunication(cellchat, min.cells = 10)
  
  # Infer the cell-cell communication at signaling pathway level
  cellchat <- computeCommunProbPathway(cellchat)
  
  #Calculate aggregate cell-cell communication network
  cellchat <- aggregateNet(cellchat)
  
  saveRDS(cellchat, '241201_cellchat_pLLg_pLLP.rds')
}

```

# Visualize output

## Overall network
```{r}
groupSize <- as.numeric(table(cellchat@idents))

par(mfrow = c(1,2), xpd = TRUE)

netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = TRUE, label.edge = FALSE, title.name = 'Number of interactions')

netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = TRUE, label.edge = FALSE, title.name = 'Interaction weights/strength')

```

### Network by cluster

```{r}
mat <- cellchat@net$weight
par(mfrow = c(2,3), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, label.edge = TRUE, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```

## Network heatmap

```{r, fig.asp = 1}
bubble <- netVisual_bubble(cellchat, sources.use = c(2), targets.use = c(1), signaling = c("NCAM", "Teneurin-latrophilin", "CDH", "CADM", "NGL", "EPHB", "L1CAM", "NEGR", "SEMA6", "CNTN", "GRN", "MPZ", "NGF", "GDNF"), remove.isolate = FALSE)

bubble_plot <- ggsave(bubble,  filename = "bubble_plot.jpeg", dpi = 600, scale = 0.25)

# Use netvisual_bubble function to pull data with "return.data" argument

bubble_plot <- netVisual_bubble(cellchat,
                 sources.use = c(2),
                 targets.use = c(1), 
                 signaling = c("NCAM", "Teneurin-latrophilin", "CDH", "CADM", "NGL", "EPHB", "L1CAM", "NEGR", "SEMA6", "CNTN", "GRN", "MPZ", "NGF", "GDNF"),
                 return.data = TRUE)

# Arrange bubble data by probability
bubble_dat <- bubble_plot$communication %>%
  arrange(desc(prob))

# Generate bubbleplot with custom order

netVisual_bubble(cellchat,
                 sources.use = c(2),
                 targets.use = c(1), 
                 pairLR.use = data.frame(interaction_name = bubble_dat$interaction_name))

# Reversed order
bubble <- netVisual_bubble(cellchat,
                 sources.use = c(2),
                 targets.use = c(1), 
                  font.size = 12,
                 pairLR.use = data.frame(interaction_name = rev(bubble_dat$interaction_name)))

ordered_bubble_plot <- ggsave(bubble,  filename = "ordered_bubble_plot.jpeg", dpi = 600)
  

```

## Chord plots (focus on cluster 1 & 2)

```{r}
netVisual_chord_gene(cellchat, sources.use = c(2), targets.use = c(1), lab.cex = 1, legend.pos.y = 30)
```

# Systems analysis of interaction

## centrality

```{r}
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP") # the slot 'netP' means the inferred intercellular communication network of signaling pathways
# Visualize the computed centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups
netAnalysis_signalingRole_network(cellchat, width = 8, height = 2.5, font.size = 10)
```

## Visualize dominant senders and receivers

```{r}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
gg1 <- netAnalysis_signalingRole_scatter(cellchat)
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
# Signaling role analysis on the cell-cell communication networks of interest
gg2 <- netAnalysis_signalingRole_scatter(cellchat, signaling = c('Teneurin-latrophilin', "MK", "LAMININ","AGRN", "NGL", "EPHB", "THBS", "CDH", "PTN", "IGF", "ANGPTL", "TENASCIN", 
         "SEMA3", "BMP", "EPHA", "CXCL", "FGF", "NOTCH", "GAS", "MIF", "PROS", "TRAIL"))
#> Signaling role analysis on the cell-cell communication network from user's input

gg1

gg2
```

## Identify dominant incoming/outgoing signals

```{r, fig.asp = 1}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", width = 5)

ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", width = 5)

ht1

ht2

ht1+ht2

outgoing <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", width = 5, signaling = c("NCAM", "Teneurin-latrophilin", "CDH", "CADM", "NGL", "EPHB", "L1CAM", "NEGR", "SEMA6", "CNTN", "GRN", "MPZ", "NGF", "GDNF"))



incoming <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", width = 5, signaling = c("NCAM", "Teneurin-latrophilin", "CDH", "CADM", "NGL", "EPHB", "L1CAM", "NEGR", "SEMA6", "CNTN", "GRN", "MPZ", "NGF", "GDNF"))



outgoing+incoming


```
## Identify patterns of signal

### load libraries
```{r}
library(NMF)
library(ggalluvial)
```

### Identify outgoing patterns

```{r}
selectK(cellchat, pattern = 'outgoing')

nPatterns <- 3
```

```{r}
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "outgoing", k = nPatterns)

netAnalysis_dot(cellchat, pattern = 'outgoing')
```


### Identify incoming patterns

```{r}
selectK(cellchat, pattern = 'incoming')

nPatterns <- 3
```

```{r}
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "incoming", k = nPatterns)

netAnalysis_dot(cellchat, pattern = 'incoming')
```

# Revisit pathways of interest

## Set pathways of interest
```{r}
poi <- c("NCAM", "Teneurin-latrophilin", "CDH", "CADM", "NGL", "EPHB", "L1CAM", "NEGR", "SEMA6", "CNTN", "GRN", "MPZ", "NGF", "GDNF")
```

## Table of every included interaction

```{r}
CellChatDB2$interaction %>%
  filter(pathway_name %in% poi) %>%
  DT::datatable()
```


## Go through pathway by pathway



```{r, fig.asp = 1, eval = FALSE}

for(i in poi){
  na1 <- netAnalysis_signalingRole_network(cellchat, signaling = i , width = 8, height = 2.5, font.size = 10)
   print(na1)
   
  np1 <- netVisual_heatmap(cellchat, signaling = i)
  print(np1)

  nac1 <- netAnalysis_contribution(cellchat, signaling = i)
  print(nac1)
  
  nb1 <- netVisual_bubble(cellchat, signaling = i, title.name = i)
  print(nb1)
  
  vp1 <- plotGeneExpression(cellchat, signaling = i, type = "violin")
  print(vp1)
}


```

# sessionInfo()

```{r}
sessionInfo()
```

