---
title: "nechiporuk_10x_analysis"
author: "nlc", "avn"
date: "4/27/2020"
output:   
  html_document: 
    toc: yes
    toc_depth: 2
    toc_float: yes
    toc_collapsed: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.align = 'center', out.width = '90%')
```

<style>
    body .main-container {
        max-width: 2160px;
    }
</style>

# Set up

## Load libraries
```{r}
library(tidyverse)
library(Seurat)
library(cowplot)
library(patchwork)
library(MAST)
library(ggrepel)
library(monocle3)
library(SeuratWrappers)
library(Matrix)
library(pheatmap)
library(RColorBrewer)



```

## Load data
```{r}
#14h data set - WT_1
data_folder1 <- "WT_14h"
WT14_data <- Read10X(data.dir = data_folder1)
WT14 <- CreateSeuratObject(counts = WT14_data,
                           project = "WT14", min.cells = 3, min.features = 200)

#18h data set 
data_folder2 <- "WT_18h"
WT18_data <- Read10X(data.dir = data_folder2)
WT18 <- CreateSeuratObject(counts = WT18_data,
                          project = "WT18", min.cells = 3, min.features = 200)

# #22h data set - WT_2
data_folder2 <- "WT_22h"
WT22_data <- Read10X(data.dir = data_folder2)
WT22 <- CreateSeuratObject(counts = WT22_data,
                           project = "WT22", min.cells = 3, min.features = 200)


# #30h data set#1 - WT_3
data_folder3 <- "WT30h_1"
WT_3_data <- Read10X(data.dir = data_folder3)
WT30_1 <- CreateSeuratObject(counts = WT_3_data,
                           project = "WT30_1", min.cells = 3, min.features = 200)

# #30h data set#2 - WT_4
data_folder4 <- "WT30h_2"
WT_4_data <- Read10X(data.dir = data_folder4)
WT30_2 <- CreateSeuratObject(counts = WT_4_data,
                           project = "WT30_2", min.cells = 3, min.features = 200)

#48h data set - WT_5
data_folder5 <- "WT48h_1"
WT_5_data <- Read10X(data.dir = data_folder5)
WT48 <- CreateSeuratObject(counts = WT_5_data,
                           project = "WT48_1", min.cells = 3, min.features = 200)


```


```{r}
data.combined <- merge(x = WT14, y = c(WT18, WT22, WT30_1, WT30_2, WT48), add.cell.ids = c("WT14", "WT18", "WT22", "WT30_1", "WT30_2", "WT48"), project = "data.combined")
data.combined
data.combined[['condition']] <- str_split(data.combined@meta.data$orig.ident, pattern = '_', simplify = TRUE)[,1]
head(colnames(data.combined))
table(data.combined$orig.ident)

```


# 'Standard' Seurat analysis

## Assign mitochondrial percent

*NOTE (NLC):* zebrafish mitochondrial genes have a lowercase prefix, using the regex pattern '^mt-' appeared to work, however only 13 mitchondrial genes are present.

```{r}
data.combined[["percent.mt"]] <- PercentageFeatureSet(data.combined, pattern = "^mt-|^COX3$")
# Visualize QC metrics as a violin plot
VlnPlot(data.combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0)


```

## QC visualization

*Note (AN):* FeatureScatter is typically used to visualize feature-feature relationships, but can be used for anything calculated by the object, i.e. columns in object metadata, PC scores etc.

*Note (NLC):* zebrafish mitochondrial genes have a lowercase prefix, 



Run this chunk to manually set cutoffs:

```{r, eval = TRUE}
nCount_RNA_limits <- c(1e3, 7.5e4)

nFeature_RNA_limits <- c(1900, 9000)

percent_mt_limits <- c(0, 5)
```


```{r}

FeatureScatter(data.combined, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident")+
  geom_vline(xintercept = nCount_RNA_limits, linetype = 'dashed', size = 1.5)+
  geom_hline(yintercept = percent_mt_limits, linetype = 'dashed', size = 1.5)

FeatureScatter(data.combined, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident")+
  geom_vline(xintercept = nCount_RNA_limits, linetype = 'dashed', size = 1.5)+
  geom_hline(yintercept = nFeature_RNA_limits, linetype = 'dashed', size = 1.5)

data.combined_metadata <- data.combined@meta.data

ggplot(data.combined_metadata, aes(x = nCount_RNA, y = nFeature_RNA))+
  geom_point(alpha = 0.1)+
  geom_density2d()+
  geom_vline(xintercept = nCount_RNA_limits, linetype = 'dashed', size = 1.5)+
  geom_hline(yintercept = nFeature_RNA_limits, linetype = 'dashed', size = 1.5)+
  scale_x_log10()

ggplot(data.combined_metadata, aes(x = nCount_RNA/nFeature_RNA, y = nFeature_RNA))+
  geom_point(alpha = 0.1)+
  geom_density2d()+
  geom_hline(yintercept = nFeature_RNA_limits, linetype = 'dashed', size = 1.5)



```

## Subset data

```{r}
#Display counts before pruning
print('Cell count before pruning:')
table(data.combined@meta.data$orig.ident)

data.combined <- subset(data.combined, subset = nFeature_RNA > min(nFeature_RNA_limits) & nFeature_RNA < max(nFeature_RNA_limits) & nCount_RNA > min(nCount_RNA_limits) & nCount_RNA < max(nCount_RNA_limits) & percent.mt > min(percent_mt_limits) & percent.mt < max(percent_mt_limits))

#Display counts after pruning
print('Cell count after pruning:')
table(data.combined@meta.data$orig.ident)
```

## Ancor and normalize data
```{r}
data.combined.list <- SplitObject(data.combined, split.by = "orig.ident")

data.combined.list <- lapply(X = data.combined.list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

anchors <- FindIntegrationAnchors(object.list = data.combined.list, dims = 1:25)
data.combined <- IntegrateData(anchorset = anchors, dims = 1:25)
#saveRDS(data.combined, file = "outputs/14_18_22_30_48h.rds")
```


## Identify variable features and scale data

```{r}
# Run the standard workflow for visualization and clustering
#data.combined <- readRDS(file = "outputs/14_18_22_30_48h.rds")
DefaultAssay(data.combined) <- "integrated"
data.combined <- ScaleData(data.combined, features = row.names(data.combined), verbose = FALSE)
data.combined <- RunPCA(data.combined, npcs = 50, verbose = FALSE)

DimPlot(data.combined, reduction = 'pca')
ElbowPlot(data.combined, ndims = 50)+
  geom_vline(xintercept = 30.5)

# t-SNE and Clustering
data.combined <- RunUMAP(data.combined, reduction = "pca", dims = 1:40)
data.combined <- FindNeighbors(data.combined, reduction = "pca", dims = 1:40)
data.combined <- FindClusters(data.combined, resolution = 2)

plot_14_48h <- DimPlot(data.combined, reduction = 'umap', label = T) + NoLegend()+
  coord_equal()

ggsave("outputs/plot_14_48h.jpeg", plot_14_48h, dpi = 600)

DimPlot(data.combined, reduction = 'umap', group.by = 'orig.ident')+
  coord_equal()

```

## color blind friendly UMAP
```{r}

my_colors <- c("5c5cc2",
"#a6bc3a",
"#733fa1",
"#59c66f",
"#b43996",
"#7cb041",
"#847deb",
"#c8aa34",
"#482070",
"#4cbe84",
"#d370d1",
"#4d7f29",
"#de4782",
"#43c29e",
"#9c1a54",
"#36dee6",
"#902a0e",
'deepskyblue2',
"#d28826",
"#353f8d",
"#9ca03c",
"#81438d",
"#7cbf76",
"#6d195d",
'darkmagenta',
"#851657",
"#3a8142",
"#e66bab",
"#255719",
"#c08adc",
"#dcae57",
"#5963ad",
"#cb692b",
"#7199e0",
"#d5533b",
"#db81c4",
"#a67724",
"#9b4486",
"#7b6a24",
"#b23679",
"#cca361",
"#6b1b4b",
"#dc8e5a",
"#6f193d",
"#d16a43",
"#a14e7c",
"#8d4d17",
"#e480a8",
"#701e0f",
"#eb5573",
"#74171f",
"#e5836a",
"#72182e",
"#dd4e4f",
"#b1486e",
"#a3553a",
"#ad264f",
"#dd7a87",
"#a73830",
"#d55b75",
"#9d323b",
"#df6f70",
"#a34656",
"#c4394f")

plot_14_48h <- DimPlot(data.combined, reduction = 'umap', label = F, pt.size = 0.1,  cols = my_colors) + NoLegend()

ggsave("outputs/plot_14_48h.jpeg", plot_14_48h,  dpi = 600, width = 210,
  height = 140, units = "mm")
```


# Directed analysis


```{r}

DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c('ret', 'irx1a'), min.cutoff = 0,
  max.cutoff = NA, order = TRUE)+
  coord_equal()

FeaturePlot(data.combined, features = c("phox2bb",
                "mafba"), min.cutoff = 0,
  max.cutoff = NA, order = TRUE)+
  coord_equal()




```


## Look at cluster identity versus condition

```{r}
data.combined.meta <- data.combined@meta.data

ggplot(data.combined.meta, aes(x = seurat_clusters, fill = orig.ident))+
  geom_bar(position = 'dodge')
```

# Analysis of clusters

## 'Automatic' subset and analysis a particular cluster
```{r}

#Set the subset cluster(s) that you're interested in

subset_cluster_id <- c(17,24)
DefaultAssay(data.combined) <- "integrated"
##
##The code below is designed to run automatically once the subset ids are selected above
##

print('Running automated ~analysis~ on the following clusters:')
print(subset_cluster_id)

#Pull the subset
data.combined_subset <- subset(data.combined, subset = seurat_clusters %in% subset_cluster_id)
#stash the original cluster identities
data.combined_subset@meta.data$original_seurat_clusters <- data.combined_subset@meta.data$seurat_clusters

#Perform relatively hands-free analysis of the selected clusters
data.combined_subset <- FindVariableFeatures(data.combined_subset)
data.combined_subset <- RunPCA(data.combined_subset)
ElbowPlot(data.combined_subset)

data.combined_subset <- FindNeighbors(data.combined_subset, dims = 1:10)
data.combined_subset <- FindClusters(data.combined_subset, resolution = 0.2)
data.combined_subset <- RunUMAP(data.combined_subset, dims = 1:10)


DimPlot(data.combined_subset, label = TRUE, label.size = 10, repel = 0.1,  group.by = 'seurat_clusters')+
  ggtitle('New clusters')
subset_plot2 <- DimPlot(data.combined_subset, label = TRUE, label.size = 10, repel= 0.1, group.by = 'original_seurat_clusters')+
  ggtitle('Original cluster identities')


data.combined.subset.meta <- data.combined_subset@meta.data

ggplot(data.combined.subset.meta, aes(x = seurat_clusters, fill = orig.ident))+
  geom_bar(position = 'dodge')


#Identify anterior lateral line neurons by alcama expression
DefaultAssay(data.combined_subset) <- "RNA"
FeaturePlot(data.combined_subset, features = c('alcama'), order = T,
            min = 0.1,
            pt.size = 1)+
  coord_equal()+scale_x_reverse()

#subset progenitors and posterior lateral line neurons 
data.combined_subset2 <- subset(x = data.combined_subset, idents = c("0", "2", "3"), invert = F)
subset2_plot <- DimPlot(data.combined_subset2, label = F, label.size = 10, repel = 0.1, pt.size = 1)+
  coord_equal()+scale_x_reverse() + scale_color_manual(values = c('deepskyblue2','darkorange1','darkmagenta'))
ggsave("outputs/subset2_plot.jpeg", subset2_plot, dpi = 600)

subset2_plot_split <- DimPlot(data.combined_subset2, label = F, label.size = 10, repel = 0.1, pt.size = .5, split.by = "condition")+
  coord_equal()+scale_x_reverse() + scale_color_manual(values = c('deepskyblue2','darkorange1','darkmagenta'))
ggsave("outputs/subset2_plot_split.jpeg", subset2_plot_split, dpi = 600)

#subset 14, 18, 22 hpf
Idents(data.combined_subset2) <- "condition" 
cell_values <- c("WT14", "WT18", "WT22")
data.combined_subset3 <- subset(x = data.combined_subset2, idents = cell_values, invert = FALSE)

DimPlot(data.combined_subset3, label = F, label.size = 10, repel = 0.1, pt.size = 1)+
  coord_equal()+scale_x_reverse()

DimPlot(data.combined_subset3, label = F, label.size = 10, repel = 0.1, pt.size=1, split.by = "condition")+
  coord_equal()+scale_x_reverse()



```
## Marker expression analysis

```{r}
#data.combined_subset2 <- readRDS(file="outputs/data.combined_subset2.rds")
#neural progenitors
DefaultAssay(data.combined_subset2) <- "RNA"
neurog1 <- FeaturePlot(data.combined_subset2, features = c('neurog1'), order = T,
            min = 0.1,
            pt.size = 2)+
  coord_equal()+scale_x_reverse()
ggsave("outputs/neurog1.jpeg", neurog1, dpi = 600)

#neurons
snap25b <- FeaturePlot(data.combined_subset2, features = c('snap25b'), order = T,
            min = 0.1,
            pt.size = 2)+
  coord_equal()+scale_x_reverse()
ggsave("outputs/snap25b.jpeg", snap25b, dpi = 600)

#pioneer and follower markers
rpz5 <- FeaturePlot(data.combined_subset2, features = c("rpz5"), order = T,
            min = 0.1,
            pt.size = 2)+
  coord_equal()+scale_x_reverse()
ggsave("outputs/rpz5.jpeg", rpz5, dpi = 600)

ret <- FeaturePlot(data.combined_subset2, features = c("ret"), order = T,
            min = 0.1,
            pt.size = 2)+
  coord_equal()+scale_x_reverse()
ggsave("outputs/ret.jpeg", ret, dpi = 600)

zfhx3 <- FeaturePlot(data.combined_subset2, features = c("zfhx3"), order = T,
            min = 0.1,
            pt.size = 2)+
  coord_equal()+scale_x_reverse()
ggsave("outputs/zfhx3.jpeg", zfhx3, dpi = 600)

hoxb5a <- FeaturePlot(data.combined_subset2, features = c('hoxb5a'), order = T,
            min = 0.1,
            pt.size = 2)+
  coord_equal()+scale_x_reverse()
ggsave("outputs/hoxb5a.jpeg", hoxb5a, dpi = 600)

nr2f2 <- FeaturePlot(data.combined_subset2, features = c('nr2f2'), order = T,
            min = 0.1,
            pt.size = 2)+
  coord_equal()+scale_x_reverse()
ggsave("outputs/nr2f2.jpeg", nr2f2, dpi = 600)


#Violoin plot for RA genes
data("data.combined_subset2")
ra_genes <- c("crabp1b", "hoxb3a",  "hoxb5a",  "nr2f2",   "zfhx3" )
ra_genes_vinplot <- VlnPlot(object = data.combined_subset2, features = ra_genes, group.by = "condition") + scale_size(range = c(5, 10)) &  stat_summary(fun.y = "mean", geom='point', size = 10, colour = "black", shape = 95)

ggsave("outputs/ra_genes_vinplot.jpeg", ra_genes_vinplot, dpi = 600,  width = 250,
  height = 200, units = "mm")
```


# pseudotime analysis using Monocle for 14-48 hpf subset
```{r}

# ...1 Convert to cell_data_set object ------------------------


#pll.subset <- subset(data.combined_subset2, subset = integrated_snn_res.0.2 %in% c(0,2,3))

pllg.cds <- new_cell_data_set(expression_data = data.combined_subset2@assays$RNA@counts,
                              cell_metadata = data.combined_subset2@meta.data)

# Fix gene_name and gene_short_name to enable gene expression plots. Solution from: https://github.com/cole-trapnell-lab/monocle3/issues/438
rowData(pllg.cds)$gene_name <- rownames(pllg.cds)
rowData(pllg.cds)$gene_short_name <- rowData(pllg.cds)$gene_name


# to get counts
counts(pllg.cds)


# ...2. Cluster cells (using clustering info from seurat's UMAP)---------------------------
# let's use the clustering information have

# assign paritions
reacreate.partition <- c(rep(1,length(pllg.cds@colData@rownames)))
names(reacreate.partition) <- pllg.cds@colData@rownames
reacreate.partition <- as.factor(reacreate.partition)


pllg.cds@clusters$UMAP$partitions <- reacreate.partition

# Assign the cluster info 

list_cluster <- data.combined_subset2@active.ident
pllg.cds@clusters$UMAP$clusters <- list_cluster


# Assign UMAP coordinate - cell embeddings

pllg.cds@int_colData@listData$reducedDims$UMAP <- data.combined_subset2@reductions$umap@cell.embeddings


# plot
cluster.before.trajectory <- plot_cells(pllg.cds,
           color_cells_by = 'cluster',
           label_groups_by_cluster = FALSE,
           group_label_size = 5) +
  theme(legend.position = "right")


# ...3. Learn trajectory graph ------------------------
pllg.cds <- learn_graph(pllg.cds, use_partition = FALSE)

plot_cells(pllg.cds,
           color_cells_by = 'cluster',
           label_groups_by_cluster = FALSE,
           label_branch_points = FALSE,
           label_roots = FALSE,
           label_leaves = FALSE,
           group_label_size = 5)


# ...4. Order the cells in pseudotime -------------------

# Write function to extract an ordered vector (highest to lowest) of cell expression for a particular gene

gene_order <- function(seurat_object = NA,
                       gene_name = NA,
                       decreasing = TRUE,
                       layer = 'data',
                       assay = 'RNA'){
  if(is.na(seurat_object)){
    print('Error - Seurat object not provided')
  }
  if(is.na(gene_name)){
    print('Error - Gene name not provided')
  }
  
  if(! is.na(seurat_object) & ! is.na(gene_name)){
    data_mat <- GetAssayData(object = seurat_object,
                             layer = layer,
                             assay = assay)[gene_name,]
    data_sort <- sort(data_mat,
                      decreasing = decreasing)
  }
  
}

# Get cell expression ordered by neurod1
neurog1_order <- gene_order(seurat_object = data.combined_subset2,
                            gene_name = 'neurog1')

# Pull the cell with highest expression
highest_exp_cell <- names(neurog1_order)[1]

# Order pseudotime rooted at highest expression of neurog1
pllg.cds <- order_cells(pllg.cds, reduction_method = 'UMAP', root_cells = highest_exp_cell)


# cells ordered by monocle3 pseudotime

pseudotime(pllg.cds)
pllg.cds$monocle3_pseudotime <- pseudotime(pllg.cds)
data.pseudo <- as.data.frame(colData(pllg.cds))

ggplot(data.pseudo, aes(monocle3_pseudotime, reorder(condition, monocle3_pseudotime, median), fill = condition)) +
  geom_violin()+ geom_point(position = position_jitter(seed = .5, width = 0.1))


# visualizing pseudotime in seurat

data.combined_subset2$pseudotime <- pseudotime(pllg.cds)
data.combined.subset2.meta <- data.combined_subset2@meta.data
pseudotime <- FeaturePlot(data.combined_subset2, features = "pseudotime", label = F)+
  coord_equal() + scale_x_reverse() & scale_color_viridis_c() 
ggsave("outputs/pseudotime.jpeg", pseudotime, dpi = 600)

#save seurat subset2 object
#saveRDS(data.combined_subset2,                   file="outputs/data.combined_subset2.rds")

```

# pseudotime analysis using Monocle for 14-22 hpf subset
```{r}

# ...1 Convert to cell_data_set object ------------------------


#pll.subset <- subset(data.combined_subset2, subset = integrated_snn_res.0.2 %in% c(0,2,3))

pllg2.cds <- new_cell_data_set(expression_data = data.combined_subset3@assays$RNA@counts,
                              cell_metadata = data.combined_subset3@meta.data)

# Fix gene_name and gene_short_name to enable gene expression plots. Solution from: https://github.com/cole-trapnell-lab/monocle3/issues/438
rowData(pllg2.cds)$gene_name <- rownames(pllg2.cds)
rowData(pllg2.cds)$gene_short_name <- rowData(pllg2.cds)$gene_name


# to get counts
counts(pllg2.cds)


# ...2. Cluster cells (using clustering info from seurat's UMAP)---------------------------
# let's use the clustering information have

# assign paritions
reacreate.partition <- c(rep(1,length(pllg2.cds@colData@rownames)))
names(reacreate.partition) <- pllg2.cds@colData@rownames
reacreate.partition <- as.factor(reacreate.partition)


pllg2.cds@clusters$UMAP$partitions <- reacreate.partition

# Assign the cluster info 

list_cluster <- data.combined_subset3@active.ident
pllg2.cds@clusters$UMAP$clusters <- list_cluster


# Assign UMAP coordinate - cell embeddings

pllg2.cds@int_colData@listData$reducedDims$UMAP <- data.combined_subset3@reductions$umap@cell.embeddings


# plot
cluster.before.trajectory <- plot_cells(pllg2.cds,
           color_cells_by = 'cluster',
           label_groups_by_cluster = FALSE,
           group_label_size = 5) +
  theme(legend.position = "right")


# ...3. Learn trajectory graph ------------------------
pllg2.cds <- learn_graph(pllg2.cds, use_partition = FALSE)

plot_cells(pllg2.cds,
           color_cells_by = 'cluster',
           label_groups_by_cluster = FALSE,
           label_branch_points = FALSE,
           label_roots = FALSE,
           label_leaves = FALSE,
           group_label_size = 5)


# ...4. Order the cells in pseudotime -------------------

# Write function to extract an ordered vector (highest to lowest) of cell expression for a particular gene

gene_order <- function(seurat_object = NA,
                       gene_name = NA,
                       decreasing = TRUE,
                       layer = 'data',
                       assay = 'RNA'){
  if(is.na(seurat_object)){
    print('Error - Seurat object not provided')
  }
  if(is.na(gene_name)){
    print('Error - Gene name not provided')
  }
  
  if(! is.na(seurat_object) & ! is.na(gene_name)){
    data_mat <- GetAssayData(object = seurat_object,
                             layer = layer,
                             assay = assay)[gene_name,]
    data_sort <- sort(data_mat,
                      decreasing = decreasing)
  }
  
}

# Get cell expression ordered by neurod1
neurog1_order <- gene_order(seurat_object = data.combined_subset3,
                            gene_name = 'neurog1')

# Pull the cell with highest expression
highest_exp_cell <- names(neurog1_order)[1]

# Order pseudotime rooted at highest expression of neurog1
pllg2.cds <- order_cells(pllg2.cds, reduction_method = 'UMAP', root_cells = highest_exp_cell)


# cells ordered by monocle3 pseudotime

pseudotime(pllg2.cds)
pllg2.cds$monocle3_pseudotime <- pseudotime(pllg2.cds)
data.pseudo <- as.data.frame(colData(pllg2.cds))

ggplot(data.pseudo, aes(monocle3_pseudotime, reorder(condition, monocle3_pseudotime, median), fill = condition)) +
  geom_violin()


# visualizing pseudotime in seurat

data.combined_subset3$pseudotime <- pseudotime(pllg2.cds)
data.combined.subset3.meta <- data.combined_subset3@meta.data
pseudotime2 <- FeaturePlot(data.combined_subset3, features = "pseudotime", label = F)+
  coord_equal() + scale_x_reverse() & scale_color_viridis_c() 
ggsave("outputs/pseudotime2.jpeg", pseudotime2, dpi = 600)

#save seurat subset3 object
#saveRDS(data.combined_subset2,                   file="outputs/data.combined_subset2.rds")

key_genes <- c("neurog1", "snap25b", "zfhx3", "nr2f2", "ret", "rpz5")
key_genes <- factor(key_genes, levels = key_genes)
subset_pllg.cds <- pllg2.cds[rowData(pllg2.cds)$gene_short_name %in% key_genes,]
pseudotime_plots <- plot_genes_in_pseudotime(subset_pllg.cds,
                         color_cells_by = 'seurat_clusters',
                         min_expr = 0.5)


```

##co-expression for 14-22 hpf
```{r}
#co-expression
VlnPlot(object = data.combined_subset3, features = c('ntrk1', "nr2f2"), slot = 'counts',split.by = 'condition') & geom_hline(yintercept = 3, size = 1)

VlnPlot(object = data.combined_subset3, features = c("ret", "zfhx3"), slot = 'counts', split.by = 'condition') & geom_hline(yintercept = 3, size = 1) 

ntrk1_nr2f2 <- FeaturePlot(data.combined_subset3,
            features = c('ntrk1', 'nr2f2'),
            blend = TRUE,
            order = T,
            min.cutoff = 0,
            max.cutoff = 3,
            slot = 'counts',
            pt.size = 1,   cols = c('gray60','yellow','magenta')) & NoAxes() &
  coord_equal() & DarkTheme()& scale_x_reverse()
ggsave("outputs/ntrk1_nr2f2.jpeg", ntrk1_nr2f2, dpi = 600)


ret_zfhx3 <- FeaturePlot(data.combined_subset3,
            features = c('ret', 'zfhx3'),
            blend = TRUE,
            order = T,
            min.cutoff = 0,
            max.cutoff = 3,
            slot = 'counts',
            pt.size = 1,   cols = c('gray60','yellow','magenta')) & NoAxes() &
  coord_equal() & DarkTheme() & scale_x_reverse()
ggsave("outputs/ret_zfhx3.jpeg", ret_zfhx3, dpi = 600)

#cells that co-express ntrk1 and nr2f2
ntrk1.cutoff <- 3
nr2f2.cutoff <- 3
length(which(FetchData(data.combined_subset3, vars ='ntrk1',  slot = 'counts') >= ntrk1.cutoff))
length(which(FetchData(data.combined_subset3, vars ='nr2f2',  slot = 'counts') >= nr2f2.cutoff))
length(which(FetchData(data.combined_subset3, vars ='ntrk1',  slot = 'counts') >= ntrk1.cutoff & FetchData(data.combined_subset3, vars ='nr2f2',  slot = 'counts') >= nr2f2.cutoff))

ret.cutoff <- 3
zfhx3.cutoff <- 3
length(which(FetchData(data.combined_subset3, vars ='ret',  slot = 'counts') >= ret.cutoff))
length(which(FetchData(data.combined_subset3, vars ='zfhx3',  slot = 'counts') >= zfhx3.cutoff))
length(which(FetchData(data.combined_subset3, vars ='ret',  slot = 'counts') >= ret.cutoff & FetchData(data.combined_subset3, vars ='zfhx3',  slot = 'counts') >= zfhx3.cutoff))

```

## pioneer and follower signature

```{r}
DefaultAssay(data.combined_subset3) <- 'RNA'
#list of top 20 pioneer DE genes from the supplemental table
pioneers.to.plot <- c("ntrk1",
                      "ret",
                      "rpz5",
                      "gfra1a",     
"arid3c",
"adarb1a",
"ntng1a",
"ngfra" ,    
 "bnc2",
"CABZ01074363.1",
"jag1b",        
"hmx4",    
 "KHDRBS3",
"vsnl1a",        
"pacsin1b",
"ptprga", 
"cspg5a",
"atp1a1a.1",
"b3gnt7",
"ptgir",
"hoxb6b")

DotPlot(data.combined_subset3, assay = "RNA", features = pioneers.to.plot, dot.scale = 10, col.min = 0, group.by = "condition") + 
    RotatedAxis()
#list of top 20 DE follower genes from the supplemental table
followers.to.plot <- c('zfhx3',
'ntrk3a',
'nr2f2',
'kctd12.2',
's100s',
'crabp1b',
'tpst1',
'sfrp1a',
'svila',
'ptpn13',
'kirrel3l',
'zgc:158291',
'capga',
'si:ch211-152c2.3',
'tlx3b',
'enc1',
'tbx2b',
'abhd15a',
'lrp8',
'mafba' )

DotPlot(data.combined_subset3, assay = "RNA", features = followers.to.plot, dot.scale = 10, col.min = 0, group.by = "condition") + 
    RotatedAxis()


data.combined_subset2 <- AddModuleScore(data.combined_subset2,
                  features = list(pioneers.to.plot),
                  name="pioneer_")
FeaturePlot(data.combined_subset2,
            features = "pioneer_1", label = F, repel = TRUE,
            pt.size = 3)+
            scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))

data.combined_subset2 <- AddModuleScore(data.combined_subset2,
                  features = list(followers.to.plot),
                  name="follower_")

FeaturePlot(data.combined_subset2,
            features = "follower_1", label = F, repel = TRUE, pt.size = 3)+
            scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))

#module score for just 14-22 hpf
data.combined_subset3 <- AddModuleScore(data.combined_subset3,
                  features = list(pioneers.to.plot),
                  name="pioneer_")
FeaturePlot(data.combined_subset3,
            features = "pioneer_1", label = F, repel = TRUE,
            pt.size = 3)+
            scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))

data.combined_subset3 <- AddModuleScore(data.combined_subset3,
                  features = list(followers.to.plot),
                  name="follower_")

FeaturePlot(data.combined_subset3,
            features = c("follower_1"), label = F, repel = TRUE, pt.size = 3)+
            scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))

FeaturePlot(data.combined_subset3,
            features = c("pioneer_1", "follower_1"), label = F, repel = TRUE, pt.size = 3, keep.scale = "all") &
            scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))&coord_equal()

gene_signature <- FeaturePlot(data.combined_subset3,
            features = c("pioneer_1", "follower_1"), label = F, repel = TRUE,
            pt.size = 1, keep.scale = "all", cols = brewer.pal(n = 9, name = "BuPu"))+coord_equal() & scale_x_reverse()
ggsave("outputs/gene_signature.jpeg", gene_signature, dpi = 600)

FeaturePlot(data.combined_subset2,
            features = c("pioneer_1", "follower_1"), label = F, repel = TRUE,
            pt.size = 2, keep.scale = "all", cols = brewer.pal(n = 9, name = "BuPu"))+coord_equal() & scale_x_reverse()

#VlnPlot(data.combined_subset2, features = c("pioneer_1", "follower_1"), split.by = "condition", same.y.lims =1)


```


## Gene expression versus pseudotime plots
```{r}


key_genes <- c("neurog1", "snap25b", "zfhx3", "nr2f2", "ret", "rpz5")
key_genes <- factor(key_genes, levels = key_genes)
subset_pllg.cds <- pllg.cds[rowData(pllg.cds)$gene_short_name %in% key_genes,]
pseudotime_plots <- plot_genes_in_pseudotime(subset_pllg.cds,
                         color_cells_by = 'integrated_snn_res.0.2',
                         min_expr = 0.5) + scale_color_manual(values = c('deepskyblue2','darkorange1','darkmagenta')) 
ggsave("outputs/pseudotime_plots.jpeg", pseudotime_plots, dpi = 600)

```


# SessionInfo
```{r}
sessionInfo()
```





