---
title: "wt-30hpf-only"
author: "Lauren Miller"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: cerulean
    toc: TRUE
    toc_float: TRUE
    toc_depth: 5
    number_sections: FALSE
    df_print: kable
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE, comment = NA, warnings = FALSE, errors = FALSE, messages = FALSE)
```

# Set up

## Load libraries
```{r}
library(tidyverse)
library(Seurat)
library(cowplot)
library(patchwork)
library(MAST)
library(dittoSeq)
library(viridis)
```

## Load data
```{r}
data_folder1 <- "WT2_new"

WT2_data <- Read10X(data.dir = data_folder1)

WT2 <- CreateSeuratObject(counts = WT2_data,
                           project = "WT_2", min.cells = 3, min.features = 200)

data_folder2 <- "WT3_new"

WT3_data <- Read10X(data.dir = data_folder2)

WT3 <- CreateSeuratObject(counts = WT3_data,
                           project = "WT_3", min.cells = 3, min.features = 200)
```

```{r}
data.combined <- merge(x = WT2, y = c(WT3), add.cell.ids = c("WT_2","WT_3"), project = "data.combined")
data.combined
data.combined[['condition']] <- str_split(data.combined@meta.data$orig.ident, pattern = '_', simplify = TRUE)[,1]
head(colnames(data.combined))
table(data.combined$orig.ident)
```



# 'Standard' Seurat analysis

## Assign mitochondrial percent

*NOTE (NLC):* zebrafish mitochondrial genes have a lowercase prefix, using the regex pattern '^mt-' appeared to work, however only 13 mitchondrial genes are present.

```{r}
data.combined[["percent.mt"]] <- PercentageFeatureSet(data.combined, pattern = "^mt-")
# Visualize QC metrics as a violin plot
VlnPlot(data.combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0)
```


## QC visualization

*Note (AN):* FeatureScatter is typically used to visualize feature-feature relationships, but can be used for anything calculated by the object, i.e. columns in object metadata, PC scores etc.

*Note (NLC):* zebrafish mitochondrial genes have a lowercase prefix, 

Run this chunk to automatically set cutoffs (set eval = TRUE in chunk parameter header):
```{r, eval = FALSE}
nCount_RNA_limits <- c(mean(data.combined@meta.data$nCount_RNA)-sd(data.combined@meta.data$nCount_RNA), mean(data.combined@meta.data$nCount_RNA)+sd(data.combined@meta.data$nCount_RNA))

nFeature_RNA_limits <- c(mean(data.combined@meta.data$nFeature_RNA)-sd(data.combined@meta.data$nFeature_RNA), mean(data.combined@meta.data$nFeature_RNA)+sd(data.combined@meta.data$nFeature_RNA))

percent_mt_limits <- c(0, mean(data.combined@meta.data$percent.mt)+sd(data.combined@meta.data$percent.mt))

print(paste0('Automatically setting count threshold to cells between ', round(nCount_RNA_limits[1]), ' and ', round(nCount_RNA_limits[2]), ' total counts.'))

print(paste0('Automatically setting feature threshold to cells between ', round(nFeature_RNA_limits[1]), ' and ', round(nFeature_RNA_limits[2]), ' total features.'))

print(paste0('Automatically setting percent.mt threshold to cells between ', round(percent_mt_limits[1]), ' and ', round(percent_mt_limits[2], 1), ' percent'))
```


## Run this chunk to manually set QC cutoffs:
```{r, eval = TRUE}
nCount_RNA_limits <- c(1e3, 7.5e4)

nFeature_RNA_limits <- c(1900, 9000)

percent_mt_limits <- c(0, 5)
```

## Visualize QC metrics before pruning
```{r}
FeatureScatter(data.combined, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident")+
  geom_vline(xintercept = nCount_RNA_limits, linetype = 'dashed', size = 1.5)+
  geom_hline(yintercept = percent_mt_limits, linetype = 'dashed', size = 1.5)

FeatureScatter(data.combined, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident")+
  geom_vline(xintercept = nCount_RNA_limits, linetype = 'dashed', size = 1.5)+
  geom_hline(yintercept = nFeature_RNA_limits, linetype = 'dashed', size = 1.5)

data.combined_metadata <- data.combined@meta.data

ggplot(data.combined_metadata, aes(x = nCount_RNA, y = nFeature_RNA))+
  geom_point(alpha = 0.1)+
  geom_density2d()+
  geom_vline(xintercept = nCount_RNA_limits, linetype = 'dashed', size = 1.5)+
  geom_hline(yintercept = nFeature_RNA_limits, linetype = 'dashed', size = 1.5)+
  scale_x_log10()

ggplot(data.combined_metadata, aes(x = nCount_RNA/nFeature_RNA, y = nFeature_RNA))+
  geom_point(alpha = 0.1)+
  geom_density2d()+
  geom_hline(yintercept = nFeature_RNA_limits, linetype = 'dashed', size = 1.5)
```


## Subset data
```{r}
#Display counts before pruning
print('Cell count before pruning:')
table(data.combined@meta.data$orig.ident)

data.combined <- subset(data.combined, subset = nFeature_RNA > min(nFeature_RNA_limits) & nFeature_RNA < max(nFeature_RNA_limits) & nCount_RNA > min(nCount_RNA_limits) & nCount_RNA < max(nCount_RNA_limits) & percent.mt > min(percent_mt_limits) & percent.mt < max(percent_mt_limits))

#Display counts after pruning
print('Cell count after pruning:')
table(data.combined@meta.data$orig.ident)
```

## Anchor data and normalize
```{r}
data.combined.list <- SplitObject(data.combined, split.by = "orig.ident")

data.combined.list <- lapply(X = data.combined.list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

anchors <- FindIntegrationAnchors(object.list = data.combined.list, dims = 1:20)

data.combined <- IntegrateData(anchorset = anchors, dims = 1:20)
```

## Normalize, identify variable features and scale data
```{r}
DefaultAssay(data.combined) <- "integrated"
data.combined <- ScaleData(data.combined, features = row.names(data.combined), verbose = FALSE)
data.combined <- RunPCA(data.combined, npcs = 30, verbose = FALSE)

DimPlot(data.combined, reduction = 'pca')
ElbowPlot(data.combined, ndims = 30)+
  geom_vline(xintercept = 10.5)

data.combined <- FindNeighbors(data.combined, reduction = "pca", dims = 1:23)
data.combined <- RunUMAP(data.combined, reduction = "pca", dims = 1:23)
data.combined <- FindClusters(data.combined, resolution = 1)


DimPlot(data.combined, reduction = 'umap', group.by = "orig.ident") + coord_equal()

DimPlot(data.combined, reduction = 'umap', label = TRUE, label.size = 5) + coord_equal()
```

## Supp. Figure 1: clusters
```{r}
mycols <- c(
"mediumvioletred"	,
"royalblue4"	,
"royalblue4"	,
"chartreuse2"	,
"goldenrod1"	,

"deepskyblue1"	,
"mediumorchid4"	,
"deepskyblue1"	,
"coral2"	,
"cyan4"	,

"goldenrod1"	,
"cyan4"	,
"mediumvioletred"	,
"mediumvioletred"	,
"cyan4"	,

"deepskyblue1"	,
"mediumorchid4"	,
"mediumorchid4"	,
"mediumvioletred"	,
"coral2"	,

"mediumseagreen"	,
"deepskyblue1"	,
"chartreuse2"	,
"coral2"	,
"mediumvioletred"	,

"chartreuse2"	,
"chartreuse2"	,
"deepskyblue1"	,
"cyan4"	,
"chartreuse2"	,

"chartreuse2"	,
"mediumvioletred", 
"royalblue4",
"chartreuse2",
"coral2"
)

all_30hpf <- DimPlot(data.combined, 
               reduction = 'umap', 
               label = T, 
               label.size = 6,
               pt.size = 0.5,
               cols = mycols) &
  NoLegend() &
  coord_equal() &
  theme(text = element_text(family = "Arial"),
          axis.title = element_text(size = 20),
          axis.text = element_text(size = 20)) &
  labs(title = "",
       x = "UMAP 1",
       y = "UMAP 2") &
  scale_x_continuous(expand = c(0.05, 0.05)) &
  scale_y_continuous(expand = c(0.05, 0.05))

all_30hpf
ggsave("all_30hpf.png", all_30hpf, dpi = 600)
```

## Figure 1: cluster identification
#### blues: cranial ganglia 
#### greens: brain and spinal cord 
#### reds: retina
#### oranges: olfactory
#### purples: pancreas
```{r}
mycols_coded <- c(
"darkcyan"	,            
"royalblue4"	,        
"seagreen"	,        
"turquoise2"	,              
"deepskyblue1"	,          

"deepskyblue1"	,        
"seagreen"	,          
"orange"	,            
"greenyellow"	,            
"tomato4"	,          

"limegreen"	,           
"violetred4"	,             
"limegreen"	,       
"darkorchid3"	,             
"darkturquoise"	,         

"coral1"	,       
"royalblue4"	,   
"plum2"	,             
"lightskyblue4"	,         
"violetred"	,           

"forestgreen"	,           
"springgreen3"	,         
"tomato4"	,         
"coral1"	,         
"steelblue3"	,       

"royalblue4"	,         
"darkcyan"	,             
"darkturquoise"	,       
"slateblue4"	,               
"darkolivegreen"	,       

"darkseagreen4"	,         
"steelblue3"	,         
"sienna3"	,             
"deepskyblue1"	,       
"darkseagreen4"	        
)

all_coded <- DimPlot(data.combined, 
                     reduction = 'umap', 
                     label = F, 
                     pt.size = 0.5, 
                     cols = mycols_coded) & 
  NoLegend() &
  coord_equal() &
  theme(text = element_text(family = "Arial"),
          axis.title = element_text(size = 20),
          axis.text = element_text(size = 20)) &
  labs(title = "",
       x = "UMAP 1",
       y = "UMAP 2") &
  scale_x_continuous(expand = c(0.05, 0.05)) &
  scale_y_continuous(expand = c(0.05, 0.05))

all_coded
ggsave("all_coded.png", all_coded, dpi = 600)
```

## Figure 1: general pLL markers
```{r}
ret <- FeaturePlot(data.combined, 
            features = c("ret"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 0.5) &
  NoAxes() & 
  coord_equal() &
  theme(text = element_text(family = "Arial"),
        plot.title = element_text(size = 30, 
                                  face = "italic"),
        legend.text=element_text(size = 20,
                                 face = "plain"),
        legend.key.size = unit(1.5, 'line'),
        legend.position = c(0.8, 0.25)) &
  labs(title = "ret")
ret
ggsave("ret.png", ret, dpi = 600)
  

rspo2 <- FeaturePlot(data.combined, 
            features = c("rspo2"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 0.5) &
  NoAxes() & 
  coord_equal() &
  theme(text = element_text(family = "Arial"),
        plot.title = element_text(size = 30, 
                                  face = "italic"),
        legend.text=element_text(size = 20,
                                 face = "plain"),
        legend.key.size = unit(1.5, 'line'),
        legend.position = c(0.8, 0.25)) &
  labs(title = "rspo2")
rspo2
ggsave("rspo2.png", rspo2, dpi = 600)

  
bmper <- FeaturePlot(data.combined, 
            features = c("bmper"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 0.5) &
  NoAxes() & 
  coord_equal() &
  theme(text = element_text(family = "Arial"),
        plot.title = element_text(size = 30, 
                                  face = "italic"),
        legend.text=element_text(size = 20,
                                 face = "plain"),
        legend.key.size = unit(1.5, 'line'),
        legend.position = c(0.8, 0.25)) &
  labs(title = "bmper")
bmper
ggsave("bmper.png", bmper, dpi = 600)

  
stmn2a <- FeaturePlot(data.combined, 
            features = c("stmn2a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 0.5) &
  NoAxes() & 
  coord_equal() &
  theme(text = element_text(family = "Arial"),
        plot.title = element_text(size = 30, 
                                  face = "italic"),
        legend.text=element_text(size = 20,
                                 face = "plain"),
        legend.key.size = unit(1.5, 'line'),
        legend.position = c(0.8, 0.25)) &
  labs(title = "stmn2a")
stmn2a
ggsave("stmn2a.png", stmn2a, dpi = 600)

  
gfra1b <- FeaturePlot(data.combined,
            features = c("gfra1b"),
            min.cutoff = "q2",
            max.cutoff = "q99",
            order = T,
            pt.size = 0.5) &
  NoAxes() &
  coord_equal() &
  theme(text = element_text(family = "Arial"),
        plot.title = element_text(size = 30, 
                                  face = "italic"),
        legend.text=element_text(size = 20,
                                 face = "plain"),
        legend.key.size = unit(1.5, 'line'),
        legend.position = c(0.8, 0.25)) &
  labs(title = "gfra1b")
gfra1b
ggsave("gfra1b.png", gfra1b, dpi = 600)

```


# Figure S1: Cluster identification
```{r}
data.combined.saved <- data.combined

de_cluster_list <- list()

for(i in sort(unique(data.combined@meta.data$seurat_clusters))){
  # Status update
  print(paste('Finding DE genes for cluster:', i))
  
  # Subset to just the cluster of interest
  data.combined.subset <- subset(data.combined, subset = seurat_clusters == i)
  
  # Find the DE genes across clusters
  de_genes_subset <- FindAllMarkers(data.combined, assay = "RNA", verbose = FALSE, only.pos = TRUE, min.diff.pct = .2, test.use = "MAST")
  
  # Stash DE gene table into a list
  de_cluster_list[[paste0('cluster_', i)]] <- de_genes_subset
}

data.combined.saved@meta.data$condition_subcluster <- paste(data.combined.saved@meta.data$seurat_clusters)


markers.to.plot_all <- c(
"myod1"	,
"myot"	,
"cav3"	,
"chrnd"	,
"col1a1b"	,
"cldni"	,
"pfn1"	,
"ecrg4b"	,
"cdx1b"	,
"lamc2"	,
"cdx4"	,
"sst2"	,
"p2ry1"	,
"trpc4a"	,
"kitb"	,
"p2rx3b"	,
"entpd3"	,
"vegfd"	,
"pcdh9"	,
"irx1a"	,
"sgk494a"	,
"pmp22b"	,
"trarg1b"	,
"cart3"	,
"LO018508.1"	,
"slc32a1"	,
"gad1a"	,
"hoxc11a"	,
"syt6a"	,
"gsnb"	,
"oclna"	,
"otx1"	,
"sp8a"	,
"emx3"	,
"dlx4b"	,
"mef2cb"	,
"bcl11aa"	,
"lhx9"	,
"tbr1b"	,
"sv2c"	,
"slc18a3a"	,
"mnx1"	,
"chata"	,
"atoh7"	,
"foxg1b"	,
"ahnak"	,
"postnb"	,
"pmp22a" ,
"krt18a.1" ,
"nkx6.2"	,
"gck"	,
"pdx1"	,
"ins"	,
"alcama",
"hmx4"	,
"vgll4l"	,
"kitlgb"	,
"krt97"	,
"slc17a7b"	,
"gfra1b"	,
"trpv1"	,
"hoxb5b"	,
"bmper"	,
"rspo2"	,
"slc7a7"	,
"fev"	,
"arxa"	,
"gcga"	,
"gad1b"	,
"lhx1a"	,
"slc6a5"	,
"nrxn3a"	,
"vsx2"	,
"rorb"	,
"rx1"	,
"rx2"	,
"spint2"	,
"nxnl2"	,
"anxa1a"	,
"rasgrf2b"	,
"hoxb9a"	,
"hoxa9a"	,
"hoxc6b"	,
"hoxd3a"	,
"hoxc3a"	,
"lhx2a"	,
"cldnb"	,
"tnks1bp1"	,
"pvalb5"	,
"scn8aa"	,
"sall1a"	,
"spock3"	,
"ntrk1"	,
"tlx2"	,
"irx2a"	,
"phox2bb"	,
"tfap2a"	,
"drgx"	,
"prdm12a"	,
"ntrk3a"	,
"shox"	,
"mab21l2"	,
"pou3f1"	,
"rfx4"	,
"sox19a"	,
"plp1a"	,
"nr2f1b"	,
"tnn"	,
"irx1b"	,
"pax2a"	,
"pvalb6" ,
"mki67" ,
"pcna"
)


bigdotplot <- DotPlot(data.combined.saved, 
        assay = "RNA", 
        features = markers.to.plot_all, 
        dot.scale = 3.5,
        col.min = 0) & 
  coord_flip() &
  theme(text = element_text(family = "Arial", 
                            color = "black"),
       axis.title.x = element_blank(),
       axis.text.x = element_text(family = "Arial",
                                  color = "black",
                                  size = 10),
       axis.line.x.bottom = element_line(size = 0.5,
                                         color = "#4e5355"),
       axis.ticks.length = unit(0.1,"cm"),
       axis.ticks = element_line(size = 0.5,
                                 color = "#4e5355"),
       axis.title.y = element_blank(),
       axis.text.y = element_text(family = "Arial",
                                  color = "black",
                                  face = "italic",
                                  size = 9.5),
       axis.line.y.left = element_line(size = 0.5,
                                         color = "#4e5355"),
       legend.title = element_text(size = 10,
                                face = "plain"),
       legend.text = element_text(size = 10,
                                face = "plain"),
       legend.key.width = unit(0.8, 'cm'),
       legend.key.height = unit(0.5, 'cm'),
       legend.position = "bottom",
       legend.justification = "center") &
  guides(colour = guide_colourbar(title = "Average Expression",
                                  title.position="top", 
                                  title.hjust = 0.5),
         size = guide_legend(title = "Percent Expressed",
                             title.position="top", 
                             title.vjust = 0.5,
                             title.hjust = 0.5,
                             label.postion = "bottom")) &
  geom_point(aes(size = pct.exp), 
             shape = 21, 
             stroke = 0.5, 
             color = "#4e5355", 
              fill = alpha("#FFFFFF", 0.0)) 

bigdotplot
ggsave("bigdotplot.png",  bigdotplot, dpi = 600)
```




# Cluster identification

## proliferating and mature neurons
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, 
            features = c("mki67", "pcna"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal() 

FeaturePlot(data.combined, 
            features = c("snap25a", "snap25b"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal() 
```

## lateral line 
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c("ret", "gfra1b"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()

#alcama marks anterior LL neurons
FeaturePlot(data.combined, features = c("alcama", "trpv1"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

## trigeminal and RBs
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c("isl2a", "kitb"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()

FeaturePlot(data.combined, features = c("ntrk2a", "ntrk3a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

## endocrine pancreas
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c("pax6b", "nkx2.2a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()

## pancreas alpha, beta, delta cells
FeaturePlot(data.combined, features = c("gcgb", "ins", "sst2", 
                                        "arxa", "pdx1", "cdx4"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

## motor neurons
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c("mnx1", "slc18a3a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
FeaturePlot(data.combined, features = c("hoxb9a", "ret"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

## retina 
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c("atoh7", "tbx2b"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()

FeaturePlot(data.combined, features = c("rx2", "six6b"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()

FeaturePlot(data.combined, features = c("vsx1", "otx2b"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

##brain
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c("ccne2", "dbx1a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()

FeaturePlot(data.combined, 
            features = c("dscaml1", "islr2", "celf5a"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal() 
```

##telencephalon
```{r}
FeaturePlot(data.combined, 
            features = c("foxp2", "tbr1b", "bcl11aa"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal() 
```

## epiphysis
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c("otx5", "aanat2", "arl3l1"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            ncol = 3,
            order = T) & NoAxes() & coord_equal()
```

## olfactory placode
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c("krt18b", "krt4", "ptgs2a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            ncol = 3,
            order = T) & NoAxes() & coord_equal()
```

## olfactory neurons
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c("pvalb5", "cldnh"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

##otic
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c("irx1b", "irx2a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

## epibranchial neurons
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, features = c("pax3a", "pitx2"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()

FeaturePlot(data.combined, features = c("drgx", "phox2bb"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

## spinal cord
```{r}
FeaturePlot(data.combined, features = c("hoxb9a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

## mesenchymal cells (neural crest?)
```{r}
DefaultAssay(data.combined) <- "RNA"
FeaturePlot(data.combined, 
            features = c("snai1a", "twist1a"),
            ncol = 2,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

# skin cells (basal epiderm cells?)
```{r}
FeaturePlot(data.combined, 
            features = c("cldni", "pfn1", "krt97"),
            ncol = 3,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()

FeaturePlot(data.combined, 
            features = c("col1a1b", "ecrg4b"),
            ncol = 2,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

# muscle
```{r}
FeaturePlot(data.combined, 
            features = c("myod1", "myot"),
            ncol = 2,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
FeaturePlot(data.combined, 
            features = c("cav3", "chrnd"),
            ncol = 2,
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T) & NoAxes() & coord_equal()
```

## Supp. Figure 1: Cell count per cluster
```{r}
data.combined.meta <- data.combined@meta.data

cell_counts <- ggplot(data.combined.meta, 
       aes(x = seurat_clusters, 
           fill = condition)) +
  geom_bar(position = 'dodge',
           fill = "#4e5355",
           width = 0.8) +
  theme(text = element_text(family = "Arial",
                            size = 20),
        axis.text.x = element_text(color = "black"),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        panel.background = element_rect(fill = NA)) +
  labs(title = "",
       x = "Cluster",
       y = "Count") +
 scale_y_continuous(expand=c(0,2))

cell_counts

ggsave("cell_counts.png", cell_counts, dpi = 600)

```


## Analysis of clusters

## 'Automatic' subset and analysis a particular cluster
```{r}
#Set the subset cluster(s) 
subset_cluster_id <- c(14)

print('Running automated ~analysis~ on the following clusters:')
print(subset_cluster_id)

#Pull the subset
DefaultAssay(data.combined) <- "integrated"
data.combined_subset <- subset(data.combined, 
                               subset = seurat_clusters %in% subset_cluster_id)

#Stash the original cluster identities
data.combined_subset@meta.data$original_seurat_clusters <- data.combined_subset@meta.data$seurat_clusters

#Perform relatively hands-free analysis of the selected clusters
data.combined_subset <- FindVariableFeatures(data.combined_subset)
data.combined_subset <- RunPCA(data.combined_subset)
ElbowPlot(data.combined_subset)

data.combined_subset <- FindNeighbors(data.combined_subset, dims = 1:5)
data.combined_subset <- RunUMAP(data.combined_subset, dims = 1:3)
data.combined_subset <- FindClusters(data.combined_subset, resolution = .05)

DimPlot(data.combined_subset)


# Pioneers
FeaturePlot(data.combined_subset, 
            features = c("ret"),
            pt.size = 3,
            label = T,
            label.size = 12,
            order = T,
            min.cutoff = "q2", 
            max.cutoff = "q99") &NoAxes() &coord_equal() 

# Followers
FeaturePlot(data.combined_subset, 
            features = c("ntrk3a"),
            pt.size = 3,
            label = T,
            label.size = 12, 
            order = T, 
            min.cutoff = "q2", 
            max.cutoff = "q99") &NoAxes() &coord_equal()



# Rename clusters
new.cluster.ids <- c("Followers", "Pioneers")
names(new.cluster.ids) <- levels(data.combined_subset)
data.combined_subset <- RenameIdents(data.combined_subset, new.cluster.ids)

DimPlot(data.combined_subset, 
        pt.size = 3, 
        label = T, 
        label.size = 12) &NoAxes() &coord_equal() &NoLegend()
```

## subset clusters
```{r}
mycols2 <- c(
"darkturquoise",
"maroon")

subclusters <- DimPlot(data.combined_subset, 
        pt.size = 4, 
        label = F, 
        label.size = 12, 
        cols = mycols2) &
  NoLegend() &
  coord_equal() &
  theme(text = element_text(family = "Arial"),
          axis.title = element_text(size = 20),
          axis.text = element_text(size = 20)) &
  labs(title = "",
       x = "UMAP 1",
       y = "UMAP 2") &
  scale_x_continuous(expand = c(0.05, 0.05)) &
  scale_y_continuous(expand = c(0.05, 0.05)) &
  theme(text = element_text(family = "Arial")) &
  annotate(geom = "text", 
           x = 4.5, 
           y = 2.5,
           label = "Pioneers", 
           size = 12) &
   annotate(geom = "text", 
           x = -4.2, 
           y = -2.5, 
           label = "Followers", 
           size = 12) 

subclusters
ggsave("subclusters.png", subclusters, dpi = 600)

subclusters_ret <- FeaturePlot(data.combined_subset, 
            features = c("ret"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 4) &
  NoAxes() & 
  coord_equal() &
  theme(text = element_text(family = "Arial"),
        plot.title = element_text(size = 40, 
                                  face = "italic"),
        legend.text=element_text(size = 25,
                                 face = "plain"),
        legend.key.size = unit(2, 'line'),
        legend.position = c(0.85, 0.18)) &
  labs(title = "ret") 
subclusters_ret
ggsave("subclusters_ret.png", subclusters_ret, dpi = 600)

subclusters_ntrk3a <- FeaturePlot(data.combined_subset, 
            features = c("ntrk3a"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 4) &
  NoAxes() & 
  coord_equal() &
  theme(text = element_text(family = "Arial"),
        plot.title = element_text(size = 40, 
                                  face = "italic"),
        legend.text=element_text(size = 25,
                                 face = "plain"),
        legend.key.size = unit(2, 'line'),
        legend.position = c(0.85, 0.18)) &
  labs(title = "ntrk3a") 
subclusters_ntrk3a
ggsave("subclusters_ntrk3a.png", subclusters_ntrk3a, dpi = 600)

subclusters_rpz5 <- FeaturePlot(data.combined_subset, 
            features = c("rpz5"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 4) &
  NoAxes() & 
  coord_equal() &
  theme(text = element_text(family = "Arial"),
        plot.title = element_text(size = 40, 
                                  face = "italic"),
        legend.text=element_text(size = 25,
                                 face = "plain"),
        legend.key.size = unit(2, 'line'),
        legend.position = c(0.85, 0.18)) &
  labs(title = "rpz5") 
subclusters_rpz5
ggsave("subclusters_rpz5.png", subclusters_rpz5, dpi = 600)

subclusters_nr2f2 <- FeaturePlot(data.combined_subset, 
            features = c("nr2f2"), 
            min.cutoff = "q2", 
            max.cutoff = "q99", 
            order = T,
            pt.size = 4) &
  NoAxes() & 
  coord_equal() &
  theme(text = element_text(family = "Arial"),
        plot.title = element_text(size = 40, 
                                  face = "italic"),
        legend.text=element_text(size = 25,
                                 face = "plain"),
        legend.key.size = unit(2, 'line'),
        legend.position = c(0.85, 0.18)) &
  labs(title = "nr2f2") 
subclusters_nr2f2
ggsave("subclusters_nr2f2.png", subclusters_nr2f2, dpi = 600)

```


## Supp Figure 2: DE genes between pioneers and followers
```{r}
pioneer_follower_DE_genes <- FindMarkers(
  object = data.combined_subset,
  assay = "RNA",
  ident.1 = 1,               
  ident.2 = 0,               
  min.diff.pct = 0.25,             
  thresh.use = 0.05,          
  test.use = "MAST"    
)

# make volcano plot
pioneer_follower_DE_genes$gene <- rownames(pioneer_follower_DE_genes)

DE_volcano_plot_nolabel <- ggplot(pioneer_follower_DE_genes, aes(x = avg_log2FC, y = -log10(p_val_adj))) +
  geom_point(aes(color = ifelse(abs(avg_log2FC) > 0.7, avg_log2FC, "black")), alpha = 0.5, size = 5) +
  scale_color_viridis(discrete = T, direction = -1, begin = 0.1, end = 0.9, option = "A") +  
  #geom_text_repel(data = subset(pioneer_follower_DE_genes, abs(avg_log2FC) > 0.7), aes(label = gene), size = 5) +
  labs(x = "Average Log2 Fold Change", y = "-log10(Adjusted p-value)", color = "Average Log2 Fold Change") +
  theme_classic() +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "gray") +
  ggtitle("Differentially expressed genes between pioneers and followers") +
  NoLegend() +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16),
        plot.title = element_text(size = 22)) +
  scale_x_continuous(limits = c(-3.2, 3), breaks = seq(-3, 3, by = 1))

DE_volcano_plot_nolabel

ggsave("DE_volcano_plot_nolabel.png", DE_volcano_plot_nolabel, dpi = 1200)

```


# Figure 1: pLL subset, pioneer and follower marker dotplot
```{r}
data.combined_subset.saved <- data.combined_subset

Idents(data.combined_subset) <- "condition"

de_cluster_list <- list()

for(i in sort(unique(data.combined_subset@meta.data$seurat_clusters))){
  # Status update
  print(paste('Finding DE genes for cluster:', i))
  
  # Subset to just the cluster of interest
  data.combined.subset <- subset(data.combined_subset, subset = seurat_clusters == i)
  
  # Find the DE genes across them
  # Try using "MAST" method (test.use = "MAST")

  de_genes_subset <- FindAllMarkers(data.combined_subset, assay = "RNA", verbose = FALSE, only.pos = TRUE, min.diff.pct = .2, test.use = "MAST")
  
  # Stash that DE gene table into a list
  de_cluster_list[[paste0('cluster_', i)]] <- de_genes_subset
}

data.combined_subset.saved@meta.data$condition_subcluster <- paste(data.combined_subset.saved@meta.data$seurat_clusters)

markers.to.plot <- c("stmn2a", "rspo2", "bmper", 
                     "ret", "gfra1a", "ntrk1", "rpz5", 
                     "ntrk3a", "crabp1b", "nr2f2", "hoxb5a", "tpst1", "zfhx3")


DE_dotplot <- DotPlot(data.combined_subset.saved, 
        assay = "RNA", 
        features = markers.to.plot, 
        dot.scale = 10, 
        cols = c("blue", "red"), 
        group.by =  "condition_subcluster", 
        scale = F) & 
  RotatedAxis() &
  theme(text = element_text(family = "Arial"),
       axis.title.x = element_blank(),
       axis.text.x = element_text(family = "Lato",
                                  face = "italic",
                                  size = 16),
       axis.line.x.bottom = element_line(size = 0.5),
       axis.ticks.length = unit(0.2,"cm"),
       axis.ticks = element_line(size = 0.5),
       axis.title.y = element_blank(),
       axis.text.y = element_text(family = "Arial",
                                  size = 16),
       axis.line.y.left = element_line(size = 0.5),
       legend.title = element_text(size = 16,
                                face = "plain"),
       legend.text = element_text(size = 15,
                                face = "plain"),
       legend.key.width = unit(0.8, 'cm'),
       legend.key.height = unit(0.6, 'cm'),
       legend.position = "bottom",
       legend.justification = "center") &
  guides(colour = guide_colourbar(title = "Average Expression",
                                  title.position="top", 
                                  title.hjust = 0.5),
         size = guide_legend(title = "Percent Expressed",
                             title.position="top", 
                             title.hjust = 0.5)) &
  scale_y_discrete(labels=c("1" = "Pioneers", "0" = "Followers")) &  
  geom_point(aes(size = pct.exp), 
             shape = 21, 
             stroke = 1, 
             color = "#4e5355", 
              fill = alpha("#FFFFFF", 0.0)) 
DE_dotplot
ggsave("DE_dotplot.png", DE_dotplot, dpi = 600)
```





